---
title: 'PSTAT 126 Final Project'

output: 
  pdf_document: 
    toc: false 
    number_sections: true 
    includes: null 
    latex_engine: xelatex 
    keep_tex: false 
    pandoc_args: null 
geometry: margin=0.7in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message =  FALSE)
knitr::opts_chunk$set(warning =  FALSE)
knitr::opts_chunk$set(error =  FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
dataset <- read.csv("C:/Users/sidre/OneDrive/Desktop/Rwork/PSTAT126/Diamonds Prices2022.csv")
```

:::: callout
::: {style="text-align: center"}
[**STUDENT NAME**]{style="color: blue;"}
:::

-   Miles Reynolds (A051P79)
-   Sid Revanuru (A058P90)
::::

## Part 1: Data Description and Descriptive Statistics

**1.1**\

```{r, echo = FALSE}
head(dataset)
set.seed(29)
head(sample(dataset, 5))
```
**1.2**\
```{r, echo = FALSE}
summary(dataset)
```

- X\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=x)) +
  geom_histogram() +
  theme_minimal()
```
- Y\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=y)) +
  geom_histogram() +
  theme_minimal()
```
- Z\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=z)) +
  geom_histogram() +
  theme_minimal()
```
- Price\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=price)) +
  geom_histogram() +
  theme_minimal()
```
- Table\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=table)) +
  geom_histogram() +
  theme_minimal()
```
This distribution looks like it is centered around 55-57 with a very slight right skew. It looks to be approximately normally distributed.\
- Depth\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=depth)) +
  geom_histogram() +
  theme_minimal()
```
The Depth variable is approximately normally distributed with a a mean around 61-61. Looking at it visually, there doesn't seem to be much of a skew.\
Carat\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=z)) +
  geom_histogram() +
  theme_minimal()
```
The Carat variable does not have much of a skew, but is heavily centered around 2-6 carats. Most of the data is concentrated within this range.
