---
title: 'PSTAT 126 Final Project'

output: 
  pdf_document: 
    toc: false 
    number_sections: true 
    includes: null 
    latex_engine: xelatex 
    keep_tex: false 
    pandoc_args: null 
geometry: margin=0.7in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message =  FALSE)
knitr::opts_chunk$set(warning =  FALSE)
knitr::opts_chunk$set(error =  FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
dataset <- read.csv("C:/Users/sidre/OneDrive/Desktop/Rwork/PSTAT126/Diamonds Prices2022.csv")
```

:::: callout
::: {style="text-align: center"}
[**STUDENT NAME**]{style="color: blue;"}
:::

-   Miles Reynolds (A051P79)
-   Sid Revanuru (A058P90)
::::

# Part 1: Data Description and Descriptive Statistics #

**1.1**\

**Dataset**\

```{r. echo = FALSE}
head(dataset)
```

**Random Sample**\

```{r, echo = FALSE}
set.seed(29)
head(sample(dataset, 5))
```

**1.2**\

**Summary**\
```{r, echo = FALSE}
summary(dataset)
```

- X\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=x)) +
  geom_histogram() +
  theme_minimal()
```

The distribution for "x" appears to be centered around 6 mm and slightly skewed right. 

- Y\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=y)) +
  geom_histogram() +
  theme_minimal()
```

The distribution for "y" appears to be strongly centered around the 6-8 range, displaying low variance compared to "x".

\newpage
- Z\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=z)) +
  geom_histogram() +
  theme_minimal()
```

The "z" variable is strongly centered around 4 mm, with a slight right skew. 

- Price\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=price)) +
  geom_histogram() +
  theme_minimal()
```

The variable "price" is skewed right, with a mean of a round 2.5k and having a minimum of around 300 with values ranging way up into the ten-thousands.

\newpage
- Table\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=table)) +
  geom_histogram() +
  theme_minimal()
```

The "Table" distribution looks like it is centered around 55-57 with a very slight right skew. It looks to be approximately normally distributed.\

- Depth\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=depth)) +
  geom_histogram() +
  theme_minimal()
```

The "Depth" variable is approximately normally distributed with a a mean around 61-61. Looking at it visually, there doesn't seem to be much of a skew.\

\newpage
- Carat\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=z)) +
  geom_histogram() +
  theme_minimal()
```

The "Carat" variable does not have much of a skew, but is heavily centered around 2-6 carats. Most of the data is concentrated within this range.\

- Cut\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=cut)) +
  geom_bar() +
  theme_minimal()
```

Most cuts appear to be "ideal" and there is more cuts deemed better than ideal than worse than ideal, indicating a slight skew left. 

\newpage
- Color\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=color)) +
  geom_bar() +
  theme_minimal()
```

The diamonds are fairly spread out over each of the colors with "G" accounting for most of the data set.

Clarity\
```{r, echo = FALSE, fig.width = 4, fig.height = 3}
ggplot(dataset, aes(x=clarity)) +
  geom_bar() +
  theme_minimal()
```

In terms of "clarity", most diamonds fall under the "S" category, with SI1 and VS2 accounting for most of the data set.
